Evaluating computational methods for modeling off-normal operation of gas centrifuge cascades

### 2018-05-22 theory.tex
first draft of theory

The cascade is built as an ideal cascade, with no losses in the separative
work, which corresponds to $\alpha =\beta = const$ for all
stage of the cascade, where $\alpha$ and $\beta$ respectively represent the feed
to product and the feed to tail enrichment factors.  $\alpha$ and $\beta$ can be
expressed as function of the abondance ($R$) or the enrichment ($N$) of respectively the
product ($R'$,$N'$) and the feed,($R$,$N$) and the feed and the tails ($R''$,$N''$) such as:
\begin{subequations} \label{eqs:alphabeta}
    \begin{equation} \label{eq:alpha_def}
        \alpha = \frac{R'}{R} = \frac{N'}{1-N'}\frac{1-N}{N} 
\end{equation}
\\
\begin{equation}\label{eq:beta-def}
        \beta = \frac{R}{R''} = \frac{N}{1-N}\frac{1-N''}{N''} 
\end{equation}
\end{subequations}

As detailled in \cite{avery} it is also possible to derive $\alpha$ from the
first principle, and express it as a function of the feed rate $F$ the separative performance $\delta
U(\theta)$, the cut $\theta$:

@@

From the mass conservation, $N = \theta N' + (1-\theta)N''$, and equations \eqref{eqs:alphabeta} it is
possible to express $\beta$ as a function of the feed abondance, $R$, the cut
$\theta$ and $\alpha$:

@@

Since $\alpha_{i}$ and $\beta_{i}$ remain constant, only the value of the cut,
$\theta_{i}$, changes in each stage $i$ of a cascade.  This algorithm assumes
that the corresponding separative power $\delta U$ (not re-computed) can be
achieved with the chosen centrifuge design, tuning other operationnal paramter
such as the rotation speed, the counter-current flow ratio...  Once $\theta_{i}$
is determined, it is possible to compute the product and the tail assay.

@@

\subsubsection{Case C}
The last model assumes that the tail to product enrichment factor remains
constant regardless to the feed assays. To compute the response of the cascade
one need to determine $\alpha$ and $\beta$ such as their product and
$\theta$ remain fixed. 
From equations \eqref{eqs:alphabeta} and the assay conservation equation $N =
\theta N' + (1-\theta)N''$ it is possible to express the product $N'$ as a function of
the feed assay $N$, $gamma$ and the cut $theta$ as one solution of the second
order equation \eqref{eq:gamma_p}:

@@

The only solution allowing product assay values ranging between 0 and 1 is the
following :
\begin{equation}\label{eq:model_b_sol}
    \frac{\gamma(N + \theta) - N 
            - \sqrt{\gamma^2(N^2 
                            - 2N\theta 
                            + \theta^2) 
                    - \gamma(2N^2 - 2\theta^2 
                            + 2N + 2\theta) 
                    + N^2 + 2N\theta + \theta^2 - 2N - 2\theta + 1} 
            - \theta + 1}
        {2\theta(\gamma - 1)}
\end{equation}

The last model is very similar to the model A.

>>>>

The cascade is built as an ideal cascade, with no losses in the separative work,
which corresponds to $\alpha =\beta = const$ for all stage of the cascade, where
$\alpha$ and $\beta$ respectively represent the feed to product and the feed to
tail enrichment factors.  $\alpha$ and $\beta$ can be expressed as function of
the abundance ($R$) or the enrichment ($N$) of respectively the product
($R'$,$N'$) and the feed,($R$,$N$) and the feed and the tails ($R''$,$N''$) such
as:
\begin{subequations} \label{eqs:alphabeta}
    \begin{equation} \label{eq:alpha_def}
        \alpha = \frac{R'}{R} = \frac{N'}{1-N'}\frac{1-N}{N} 
\end{equation}
\\
\begin{equation}\label{eq:beta-def}
        \beta = \frac{R}{R''} = \frac{N}{1-N}\frac{1-N''}{N''} 
\end{equation}
\end{subequations}

As detailed in \cite{avery} it is also possible to derive $\alpha$ from the
first principle, and express it as a function of the feed rate $F$ the
separative performance $\delta U(\theta)$, the cut $\theta$:

@@

From the mass conservation, $N = \theta N' + (1-\theta)N''$, and equations
\eqref{eqs:alphabeta} it is possible to express $\beta$ as a function of the
feed abundance, $R$, the cut $\theta$ and $\alpha$:

@@

Since $\alpha_{i}$ and $\beta_{i}$ remain constant, only the value of the cut,
$\theta_{i}$, changes in each stage $i$ of a cascade.  This algorithm assumes
that the corresponding separative power $\delta U$ (not re-computed) can be
achieved with the chosen centrifuge design, tuning other operational parameter
such as the rotation speed, the counter-current flow ratio...  Once $\theta_{i}$
is determined, it is possible to compute the product and the tail assay.

@@

\subsubsection{Case C}
The last model assumes that the tail to product enrichment factor remains
constant regardless to the feed assays. To compute the response of the cascade
one need to determine $\alpha$ and $\beta$ such as their product and
$\theta$ remain fixed. 
From equations \eqref{eqs:alphabeta} and the assay conservation equation $N =
\theta N' + (1-\theta)N''$ it is possible to express the product $N'$ as a function of
the feed assay $N$, $\gamma$ and the cut $\theta$ as one solution of the second
order equation \eqref{eq:gamma_p}:

@@

The only solution allowing product assay values ranging between 0 and 1 is the
following :
\begin{equation}\label{eq:model_b_sol}
    \frac{\gamma(N + \theta) - N 
            - \sqrt{\gamma^2(N^2 
                            - 2N\theta 
                            + \theta^2) 
                    - \gamma(2N^2 - 2\theta^2 
                            + 2N + 2\theta) 
                    + N^2 + 2N\theta + \theta^2 - 2N - 2\theta + 1} 
            - \theta + 1}
        {2\theta(\gamma - 1)}
\end{equation}
Once the product assay is known, one can trivially determine the tail assay,
$\alpha$ and $\beta$ using equations \eqref{eqs:alphabeta} and mass
conservation.

Similarly as model A, because the cut values remain constant, the flows don't
need to be recomputed, and the correct assays, $\alpha$ and $\beta$ are
determined through iterative blending of the product assays of the previous
stage and the tails assay of the next stage using equation
\eqref{eq:model_b_sol}.
